Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-11-09 bbbq_article.tex
Describe IEDB and 'iedbr'

To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
we used two sources of data of naturally occurring epitopes.  
The first type of data are two randomly picked 
peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
and MHC-II \cite{bergseng2015different}.
The second type of data is obtained from IEDB, 
the Immune Epitope Databse \cite{vita2019immune}.
All these sources contain sequences of naturally occurring
epitopes. 
For each of the detected epitopes,
its possible location(s) in 
a human reference proteome,
with UniProt ID UP000005640\_9606, was mapped.
For the epitopes that were present in the proteome exactly once,
the topology of the proteins in which these epitopes were located 
was predicted using both \verb;TMHMM; \cite{krogh2001predicting} 
and \verb;PureseqTM; \cite{wang2019efficient}. 
From this topology, we determined if the epitope
overlapped with a TMH.
For the second we queried IEDB for epitope sequences presented by B and T cells
in healthy humans, for the haplotypes used in this study.

@@

\begin{table}[]
  \begin{tabular}{llll}
    Goal & Tool & Reference \\ 
    \hline
    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
    Combine all                       & bbbq                      & \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
  }
  \label{table:software_used}
\end{table}

@@

To retrieve the data from the NCBI databases the
\verb;rentrez; R package \cite{rentrez} was used
that calls the NCBI website's API. To provide for a 
stable user experience for all users, 
this API limits the user to 3 calls per second.
Additionally, the API splits the result of a bigger
query into multiple pages, each of which needs one API call.
We wrote the \verb;sprentrez; package \cite{sprentrez} to provide for 
bigger queries of multiple (and delayed) API calls.

>>>>

To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
we used two sources of data of naturally occurring epitopes.  
The first type of data are two randomly picked 
peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
and MHC-II \cite{bergseng2015different}.
The second type of data is obtained from , 
the Immune Epitope Database (IEDB) \cite{vita2019immune}.
All these sources contain sequences of naturally occurring
epitopes. 
For each of the detected epitopes,
its possible location(s) in 
a human reference proteome,
with UniProt ID UP000005640\_9606, was mapped.
For the epitopes that were present in the proteome exactly once,
the topology of the proteins in which these epitopes were located 
was predicted using both \verb;TMHMM; \cite{krogh2001predicting} 
and \verb;PureseqTM; \cite{wang2019efficient}. 
From this topology, we determined if the epitope
overlapped with a TMH.
For the second we queried IEDB for epitope sequences presented by B and T cells
in healthy humans, for the haplotypes used in this study.

@@

\begin{table}[]
  \begin{tabular}{llll}
    Goal & Tool & Reference \\ 
    \hline
    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Call TMHMM from R                 & \verb;tmhmm;              & \cite{tmhmm} \\
    Call PureseqTM from R             & \verb;pureseqtmr;         & \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & \verb;netmhc2pan;         & \cite{netmhc2pan} \\
    Work with IEDB                    & \verb;iedbr;              & \cite{iedbr} \\
    Work with \verb;rentrez;          & \verb;sprentrez;          & \cite{sprentrez} \\
    Combine all                       & \verb;bbbq;               & \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
  }
  \label{table:software_used}
\end{table}

@@

To retrieve the data from the NCBI databases the
\verb;rentrez; R package \cite{rentrez} was used
that calls the NCBI database's API. 
The NCBI database provides a stable user experience for all users, 
by limiting its API to 3 calls per second per user.
Additionally, the API splits the result of a bigger
query into multiple pages, each of which needs one API call.
The \verb;sprentrez; package \cite{sprentrez} provides for 
bigger queries of multiple (and delayed) API calls.

% \paragraph{IEDB software}

To retrieve the data from the IEDB  databases \cite{vita2019immune}, 
the \verb;iedbr; R package \cite{iedbr} was written,
to calls the IEDB database's API. 
Similar to the NCBI database,
the IEDB has a limit to 1 call per second per user
and allows a query results to return 10k results maximally.
The \verb;iedbr; package \cite{iedbr} allows for bigger queries.
